### 2019-07-08 cover-letter/cover-letter.tex
Cover letter

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage{hyperref,cite}
%\signature{Daniel Antunes, Jacques-Olivier Lachaud, Hugues Talbot}
%\address{}

\begin{document}
%\begin{letter}{}
%\opening{Dear editors:}

Dear editors:\\

With respect to the work in \cite{antunes19}, the authors have included three new contributions. In section three, we describe a new local combinatorial optimization method to evolve a flow driven by the minimization of the digital elastica energy. The method is suitable for two multigrid convergent estimators of curvature, and the results are illustrated with figures and a small discussion about the running time. In section four, we extend the digital evolution model described in \cite{antunes19}. In particular, we generalize the evolution model in and we define a family of energies, each one giving rise to a new evolution model. We've included a discussion with respect to the QPBOP optimization method and how the number of unlabeled pixels varies in the family of energies. Remarkably, QPBOP labels all variables for one of the energies in the family. Finally, in section five, we compared our segmentation method with two other algorithms: GrabCut \cite{rother04grabcut} and Schoenemmans's \cite{schoenemann09linear}, and we illustrate the results with several figures.

\bibliographystyle{plain}
\bibliography{../jmiv.bib}

%\end{letter}

\end{document}