### 2019-07-08 cover-letter/cover-letter.tex
Update cover letter

\documentclass{letter}
%\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage{hyperref,cite}
\usepackage[a4paper, total={7in, 10in}]{geometry}
%\signature{Daniel Antunes, Jacques-Olivier Lachaud, Hugues Talbot}
%\address{}

\makeatletter
\newenvironment{thebibliography}[1]
     {\list{\@biblabel{\@arabic\c@enumiv}}%
           {\settowidth\labelwidth{\@biblabel{#1}}%
            \leftmargin\labelwidth
            \advance\leftmargin\labelsep
            \usecounter{enumiv}%
            \let\p@enumiv\@empty
            \renewcommand\theenumiv{\@arabic\c@enumiv}}%
      \sloppy
      \clubpenalty4000
      \@clubpenalty \clubpenalty
      \widowpenalty4000%
      \sfcode`\.\@m}
     {\def\@noitemerr
       {\@latex@warning{Empty `thebibliography' environment}}%
      \endlist}
\newcommand\newblock{\hskip .11em\@plus.33em\@minus.07em}
\makeatother

\signature{Daniel Antunes, Jacques-Olivier Lachaud, Hugues Talbot}
\begin{document}
\begin{letter}{Cover letter for paper ``An Elastica-driven Digital Curve Evolution Model for Image Segmentation''}
\opening{Dear editors,}

This paper is an extension version of the DGCI'19 paper ``Digital Curvature Evolution Model for Image Segmentation'' \cite{antunes19}. We have included three new contributions. In section three, we describe a new local combinatorial optimization method to evolve a flow driven by the minimization of the digital elastica energy. The method is suitable for two multigrid convergent estimators of curvature, and the results are illustrated with figures and a small discussion about the running time. In section four, we extend the digital evolution model described in \cite{antunes19}. In particular, we generalize the curve evolution model  and we define a family of energies, each one giving rise to a new evolution model. We've included a discussion with respect to the QPBOP optimization method and how the number of unlabeled pixels varies in the family of energies. Remarkably, QPBOP labels all variables for one of the energies in the family. Finally, in section five, we compared our segmentation method with two other algorithms: GrabCut \cite{rother04grabcut} and Schoenemmans's \cite{schoenemann09linear}, and we illustrate the results with several figures.

Thank you for your time and consideration.

\closing{Yours Faithfully,}

\bibliographystyle{plain}
\bibliography{../jmiv.bib}

\end{letter}

%\end{letter}

\end{document}