### 2020-01-07 Revision/Biographies/biographies.tex
bibliographies

\documentstyle[a4]{paper}

\begin{document}
\paragraph{Daniel Antunes}
A brilliant PhD student from Brazil.

\paragraph{Jacques-Olivier Lachaud}
A splendid researcher, great mentor and fast friend.

\paragraph{Hugues Talbot}
Hugues Talbot received the Habilitation from Universite Paris-Est in 2013, the PhD from \'Ecole des Mines de Paris in
1993 and the engineering degree from Ecole Centrale de Paris in 1989. He was principal research scientist at CSIRO,
Sydney, from 1994 to 2004. He is now a professor at CentraleSupelec, Universite Paris-Saclay and member of the Inria
OPIS team. He is the co-author of 6 books and over 200 articles in the area of mathematical morphology, discrete
geometry, combinatorial and continuous optimization.
\end{document}