### 2020-01-07 Revision/Biographies/biographies.tex
biographie daniel

\documentstyle[a4]{paper}

\begin{document}
\paragraph{Daniel Antunes}
Daniel Antunes received his B.S. in computer science from the Universidade Federal do Rio de Janeiro, Brazil (2014) and his M.S. in  computer science from Université Paris-Est, France (2016). Since February 2017 he is a Ph.D student at the LAMA laboratory at Université Savoie Mont Blanc, France. His research interest is focused on digital geometry and combinatorial optimization applied to image analysis.

\paragraph{Jacques-Olivier Lachaud}
A splendid researcher, great mentor and fast friend.

\paragraph{Hugues Talbot}
Hugues Talbot received the Habilitation from Universite Paris-Est in 2013, the PhD from \'Ecole des Mines de Paris in
1993 and the engineering degree from Ecole Centrale de Paris in 1989. He was principal research scientist at CSIRO,
Sydney, from 1994 to 2004. He is now a professor at CentraleSupelec, Universite Paris-Saclay and member of the Inria
OPIS team. He is the co-author of 6 books and over 200 articles in the area of mathematical morphology, discrete
geometry, combinatorial and continuous optimization.
\end{document}