### 2020-01-08 Revision/Biographies/biographies.tex
photos

\documentstyle[a4]{paper}

\begin{document}
\paragraph{Daniel Antunes}
Daniel Antunes received his B.S. in computer science from the Universidade Federal do Rio de Janeiro, Brazil (2014) and his M.S. in  computer science from Universit\'e Paris-Est, France (2016). Since February 2017 he is a Ph.D student in the laboratory of mathematics (LAMA) at Universit\'e Savoie Mont Blanc, France. His research interests is focused on digital geometry and combinatorial optimization applied to image analysis.

\paragraph{Jacques-Olivier Lachaud}
Jacques-Olivier Lachaud graduated from ENSIMAG engineering school in
Computer Science in 1994 and received a Ph.D. degree in computer
science from Joseph Fourier University (Grenoble, France) in 1998 and
the Habilitation from University Bordeaux 1 in 2006. He is now a
Professor of Computer Science at the University Savoie Mont Blanc
(Chambéry, France) and works in the laboratory of mathematics
(LAMA). His research interests are in image analysis and geometry
processing, more specifically deformable models, energy-minimizing
techniques, digital geometry and topology. He has written over 100
articles in international journals and conferences on these topics.
%% JO :-D A splendid researcher, great mentor and fast friend.

\paragraph{Hugues Talbot}
Hugues Talbot received the Habilitation from Universite Paris-Est in
2013, the PhD from \'Ecole des Mines de Paris in 1993 and the
engineering degree from Ecole Centrale de Paris in 1989. He was
principal research scientist at CSIRO, Sydney, from 1994 to 2004; and
a professor in computer science at ESIEE Paris from 2004 to 2018. He
is now a professor in the mathematics department at CentraleSupelec,
Universite Paris-Saclay and member of the Inria OPIS team. He is the
co-author or co-editor of 7 books and over 250 articles in the area of
mathematical morphology, discrete geometry, combinatorial and
continuous optimization.
\end{document}